How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in the Pacific Pond Turtle (\textit{Emys marmorata})

### 2015-08-26 doc/draft_turtle.tex
added short text about secondary dataset

We assigned a classification to each specimen for the different binning schemes based on geographic occurrence data recorded in museum collection archives. When precise latitude and longitude information were not available we estimated them from locality information. Because \citet{Spinks2005}, \citet{Spinks2010}, and \citet{Spinks20140} did not use vouchered specimens we were not able to example the sample individuals. Therefore our specimen classifications were based solely on the geographic information, not explicit assignment in previous studies. Because the exact barriers between different biogeographic regions are unknown and unclear, we represented each hypothesis with two schemes for a total of six different schemes. The schemes differed based on where geographic boundaries were assigned. This changes the classification of certain individuals near the boundaries between groups, providing a test of the robustness of the classification schemes.

@@

@@

@@

>>>>

We assigned a classification to each specimen for the different binning schemes based on geographic occurrence data recorded in museum collection archives. When precise latitude and longitude information were not available we estimated them from locality information. Because \citet{Spinks2005}, \citet{Spinks2010}, and \citet{Spinks2014} did not use vouchered specimens we were not able to example the sample individuals. Therefore our specimen classifications were based solely on the geographic information, not explicit assignment in previous studies. Because the exact barriers between different biogeographic regions are unknown and unclear, we represented each hypothesis with two schemes for a total of six different schemes. The schemes differed based on where geographic boundaries were assigned. This changes the classification of certain individuals near the boundaries between groups, providing a test of the robustness of the classification schemes.

@@

\subsection{Comparison with clear-cut example}
In addition to the \textit{E. marmorata/pallida} case, we also analyzed a set of seven morphologically distinct species in order to test if the approaches used here actually work. The seven species analyzed were \textit{Emys blandigii}, \textit{Terrapene coahuila}, \textit{Clemmys guttata}, \textit{Glyptemys insculpta}, \textit{Glyptemys muhlenbergii}, \textit{Emys orbicularis}, and \textit{Terrapene ornata} with a total of XXX specimens analyzed in total. Just as in the \textit{E. marmorata/pallida} example, we analyzed plastron shape variation. The support for a classification scheme with seven categories was evaluated using random forest, linear discriminate analysis, and multinomial logistic regression where the number of PCs used as predictors ranged from 2 to 11. Data was split into training and testing sets and model performance for both sets was evaluated using the AUC metric, just as above.

@@

\section{Comparison with clear-cut example}
In contrast with the above results, the analysis of the clearly distinct species SUPPLEMENT demonstrates almost perfect classification of the both the in-sample and out-of-sample datasets (Table \ref{tab:second_res}). These results demonstrate that if there were strong distinctions between the states of the classification schemes, the methods used here would be able to recover this.

@@

\begin{table}[ht]
  \centering
  \begin{tabular}{ l r r r }
    \hline
    & \# predictors & In-sample AUC & Out-of-sample AUC \\ 
    \hline
    random forest &   11 & 1.000 & 0.999 \\ 
    linear discriminate analysis &   10 & 1.000 & 1.000 \\ 
    multinomial logistic regression &    6 & 1.000 & 0.998 \\ 
    \hline
  \end{tabular}
  \caption{Results from classification model estimates of the secondary, multi-species dataset. Comparison of in-sample and out-of-sample AUC of the best performing model, along with the number of predictors. AUC values range between 0.5 (no better than random) and 1 (perfect classification).}
  \label{tab:second_res}
\end{table}
